### 2013-06-10 doc/dissertation/document/chapters/spn_equations/spn_equations.tex
adding serial verification data for the eigenvalue

%%---------------------------------------------------------------------------%%
\section{MCSA Verification}
\label{sec:spn_mcsa_verification}
Through the numerical studies presented up to this point in this
chapter, we have developed a set of preconditioning techniques that
permit MCSA to be used with the $SP_N$ form of the neutron transport
equation and applied to a difficult fuel assembly criticality
problem. Along with these preconditioning techniques, a set a varying
MCSA parameters were analyzed to find those that yielded the best
performance for this particular problem. In addition, several steps
were taken to mitigate the dense composite linear operators that arise
when performing explicit MCSA preconditioning in order to achieve
convergence. Using these results, we compare MCSA directly to
conventional linear solvers that would be typically used to solve the
$SP_N$ form of the transport equation as presented here. In
particular, we will use a pair of production Krylov solvers with which
we will compare direct numerical results in order to verify MCSA
solutions and iterative and CPU timing performance in order to gauge
feasibility and identify areas that require further research.

In order to verify the correctness of the MCSA method and
implementation used for this work, MCSA solutions to the fuel assembly
problem with 1, 2, and 4 energy groups using $SP_1$ discretization
will be directly compared over all group and cell fluxes computed by
converging the criticality problem using both BiCGStab and GMRES
production implementations from the Trilinos scientific computing
library Aztec \citep{heroux_overview_2005}. In addition to comparing
the cell and group fluxes, the number of eigenvalue iterations
required to converge the problem along with the eigenvalue itself will
be compared as a means of verification.

For the verification, all solvers will be preconditioned with ILUT
using a drop tolerance of \sn{1}{-5} and a fill level of 5. For GMRES,
no restrictions were placed on the size of the subspace and therefore
no restarts occurred. When the collision estimator was used with MCSA,
20,000 stochastic histories were used to compute the correction for
every energy group in the problem (40,000 and 80,000 histories total
for the 2 and 4 group calculations respectively) and 500 stochastic
histories used for each energy group when the expected value estimator
was used (1,000 and 1,500 histories total for the 2 and 4 group
calculations). The reduced domain approximation was also applied in
conjuction with the ILUT preconditioning using a fill level of 100 and
a threshold value of \sn{1}{-10} to reduce the density of states in
the Monte Carlo problem. For relaxation parameters, all MCSA
compuations used a Neumann relaxation parameter of 0.7 while a
Richardson relaxation of 1.1 was used with the collision estimator and
1.0 with the expected value estimator as determined by the previous
analysis. Table~\ref{tab:spn_solver_defs} gives definitions for the
solvers used to generate the results in the remainder of this
section. For the energy group structures,
Table~\ref{tab:spn_group_structure} gives the lower bounds of each
group in eV (an implicit upper bound of 20 MeV is assumed).
\begin{table}[h!]
  \begin{center}
    \begin{tabular}{ll}\hline\hline
      \multicolumn{1}{l}{\textbf{Name}} & 
      \multicolumn{1}{l}{\textbf{Definition}} \\
      BiCGStab-ILUT & BiCGStab preconditioned with ILUT \\
      GMRES-ILUT & GMRES preconditioned with ILUT \\
      MCSA-ILUT-R-C & MCSA preconditioned with ILUT using Richardson \\ 
                    & fixed point iteration and collision estimator \\
      MCSA-ILUT-MR-C & MCSA preconditioned with ILUT using MINRES \\
                     & fixed point iteration and collision estimator \\
      MCSA-ILUT-R-EV & MCSA preconditioned with ILUT using Richardson \\
                     & fixed point iteration and expected value estimator \\
      MCSA-ILUT-MR-EV & MCSA preconditioned with ILUT using MINRES \\
                      & fixed point iteration and expected value estimator \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Solver definitions used for MCSA verification and
      performance analysis.}}
  \label{tab:spn_solver_defs}
\end{table}
\begin{table}[h!]
  \begin{center}
    \begin{tabular}{cl}\hline\hline
      \multicolumn{1}{c}{\textbf{Number of Groups}} & 
      \multicolumn{1}{l}{\textbf{Lower Bounds (eV)}} \\
      1 & \{ \sn{1}{-5} \} \\
      2 & \{ \sn{1}{-1}, \sn{1}{-5} \} \\
      4 & \{ \sn{1}{1}, \sn{1}{0}, \sn{1}{-1}, \sn{1}{-5} \} \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Energy group lower bounds for the multigroup fuel
      assembly criticality problem in electron volts.}}
  \label{tab:spn_group_structure}
\end{table}

>>>>

%%---------------------------------------------------------------------------%%
\section{MCSA Verification}
\label{sec:spn_mcsa_verification}
Through the numerical studies presented in this chapter, we have
developed a set of preconditioning techniques that permit MCSA to be
used with the $SP_N$ form of the neutron transport equation and
applied to a difficult fuel assembly criticality problem. Along with
these preconditioning techniques, a set a varying MCSA parameters were
analyzed to find those that yielded the best performance for this
particular problem. In addition, several steps were taken to mitigate
the dense composite linear operators that arise when performing
explicit MCSA preconditioning in order to achieve convergence. Using
these results, we compare MCSA directly to conventional linear solvers
that would be typically used to solve the $SP_N$ form of the transport
equation as presented here. In particular, we will use a pair of
production Krylov solvers with which we will compare direct numerical
results in order to verify MCSA solutions in this section and
iterative and CPU timing performance in the next order to gauge
feasibility and identify areas that require further research.

In order to verify the correctness of the MCSA method and
implementation used for this work, MCSA solutions to the fuel assembly
problem with 1, 2, and 4 energy groups using $SP_1$ discretization
will be directly compared over all group and cell fluxes computed by
converging the criticality problem using both BiCGStab and GMRES
production implementations from the Trilinos scientific computing
library Aztec \citep{heroux_overview_2005}. In addition to comparing
the cell and group fluxes, the number of eigenvalue iterations
required to converge the problem along with the eigenvalue itself will
be compared as a means of verification. If all values computed using
MCSA agree with the results from the production Krylov solvers within
the solution tolerance then the MCSA solutions will be deemed verified
for these problems. Additionaly, this verification is only applicable
to serial results with parallel verification for this problem provided
in Chapter~\ref{ch:parallel_methods}.

For the verification calculations, all solvers will be preconditioned
with ILUT using a drop tolerance of \sn{1}{-5} and a fill level of
5. For GMRES, no restrictions were placed on the size of the subspace
and therefore no restarts occurred. When the collision estimator was
used with MCSA, \sn{2}{4} stochastic histories were used to compute
the correction for every energy group in the problem (\sn{4}{4} and
\sn{8}{4} histories total for the 2 and 4 group calculations
respectively) corresponding to approximately 1 stochastic history per
DOF. When the expected value estimator was used, \sn{5}{2} stochastic
histories used for each energy group (\sn{1}{3} and \sn{1.5}{3}
histories total for the 2 and 4 group calculations) giving
approximately 1 stochastic history for every 5 DOFs. The reduced
domain approximation was also applied in conjuction with the ILUT
preconditioning using a fill level of 100 and a threshold value of
\sn{1}{-10} to reduce the density of states in the Monte Carlo
problem. For relaxation parameters, all MCSA compuations used a
Neumann relaxation parameter of 0.7 while a Richardson relaxation of
1.1 was used with the collision estimator and 1.0 with the expected
value estimator as determined by the previous analysis of the
relaxation parameters. Table~\ref{tab:spn_solver_defs} gives
definitions for the solvers used to generate the results in the
remainder of this section. For the energy group structures,
Table~\ref{tab:spn_group_structure} gives the lower bounds of each
group in eV (an implicit upper bound of \sn{2}{6} eV is assumed for
group 0).
\begin{table}[h!]
  \begin{center}
    \begin{tabular}{ll}\hline\hline
      \multicolumn{1}{l}{\textbf{Name}} & 
      \multicolumn{1}{l}{\textbf{Definition}} \\
      BiCGStab-ILUT & BiCGStab preconditioned with ILUT \\
      GMRES-ILUT & GMRES preconditioned with ILUT \\
      MCSA-ILUT-R-C & MCSA preconditioned with ILUT using Richardson \\ 
                    & fixed point iteration and collision estimator \\
      MCSA-ILUT-MR-C & MCSA preconditioned with ILUT using MINRES \\
                     & fixed point iteration and collision estimator \\
      MCSA-ILUT-R-EV & MCSA preconditioned with ILUT using Richardson \\
                     & fixed point iteration and expected value estimator \\
      MCSA-ILUT-MR-EV & MCSA preconditioned with ILUT using MINRES \\
                      & fixed point iteration and expected value estimator \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Solver definitions used for MCSA verification and
      performance analysis.} \textit{The ILUT preconditioner
      parameters were identical for all calculations and solvers.}}
  \label{tab:spn_solver_defs}
\end{table}
\begin{table}[h!]
  \begin{center}
    \begin{tabular}{cl}\hline\hline
      \multicolumn{1}{c}{\textbf{Number of Groups}} & 
      \multicolumn{1}{l}{\textbf{Lower Bounds (eV)}} \\
      1 & \{ \sn{1}{-5} \} \\
      2 & \{ \sn{1}{-1}, \sn{1}{-5} \} \\
      4 & \{ \sn{1}{1}, \sn{1}{0}, \sn{1}{-1}, \sn{1}{-5} \} \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Energy group lower bounds for the multigroup fuel
      assembly criticality problem in electron volts.} \textit{An
      implicit upper bound of \sn{2}{6} eV is assumedfor group zero.}}
  \label{tab:spn_group_structure}
\end{table}

For every combination of energy group structure and solver, the
eigenvalue problem was converged with an eigensolver tolerance of
\sn{1}{-6} and a linear solver tolerance of \sn{1}{-8} in a serial
calculation. Table~\ref{tab:serial_ev_results} gives the eigenvalue
computed by each solver for each problem as well as the number of
eigenvalue iterations required to converge. For every problem, each
solver converged to the same eigenvalue within the tolerance of the
eigensolver in the same number of eigenvalue iterations.
\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lccc}\hline\hline
      \multicolumn{1}{c}{\textbf{Solver}} & 
      \multicolumn{1}{c}{\textbf{Number of Groups}} & 
      \multicolumn{1}{c}{\textbf{Iterations}} &
      \multicolumn{1}{c}{\textbf{Eigenvalue}} \\
      \hline
      BiCGStab-ILUT & 1 & 25 & 1.1059870 \\
      GMRES-ILUT & 1 & 25 & 1.1059870 \\
      MCSA-ILUT-R-C & 1 & 25 & 1.1059870 \\
      MCSA-ILUT-MR-C & 1 & 25 & 1.1059870 \\
      MCSA-ILUT-R-EV & 1 & 25 & 1.1059870 \\
      MCSA-ILUT-MR-EV & 1 & 25 & 1.1059870 \\
      \hline
      BiCGStab-ILUT & 2 & 35 & 1.1552390 \\
      GMRES-ILUT & 2 & 35 & 1.1552390 \\
      MCSA-ILUT-R-C & 2 & 35 & 1.1552390 \\
      MCSA-ILUT-MR-C & 2 & 35 & 1.1552390 \\
      MCSA-ILUT-R-EV & 2 & 35 & 1.1552390 \\
      MCSA-ILUT-MR-EV & 2 & 35 & 1.1552390 \\
      \hline
      BiCGStab-ILUT & 4 & 31 & 1.1758500 \\
      GMRES-ILUT & 4 & 31 & 1.1758500 \\
      MCSA-ILUT-R-C & 4 & 31 & 1.1758500 \\
      MCSA-ILUT-MR-C & 4 & 31 & 1.1758500 \\
      MCSA-ILUT-R-EV & 4 & 31 & 1.1758500 \\
      MCSA-ILUT-MR-EV & 4 & 31 & 1.1758500 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Serial eigensolver verification results for the
      fuel assembly problem.} \textit{Each calculation was converged
      with the same eigensolver and linear solver tolerances.}}
  \label{tab:serial_ev_results}
\end{table}

For each calculation, the scalar flux in each energy group was
compared over the entire spatial with the GMRES solution used as the
reference.
